Combinatorics of the geometry of Wilson loop diagrams II: Grassmann necklaces, dimensions, and denominators

### 2019-06-14 GWLDCII.tex
turned lemma 3.14 into cor 2.11

@@

We need one more lemma before we can prove that Algorithm~\ref{alg:put GN on WLD} does in fact give the Grassmann necklace of the positroid associated to $W$.

\begin{lem}\label{lem basis as perm}
Let $W = (\cP,[n])$ be an admissible Wilson loop diagram and let $M(W)$ be its associated matroid. A subset $J \subseteq [n]$ is an independent set of $M(W)$ if and only if there exists an injective set map $f : J \rightarrow \cP$ with the property that for each $j\in J$ we have $j \in V(f(j))$.
\end{lem}

In particular, this lemma says that a subset $J$ of $[n]$ is a basis of $M(W)$ if and only if there is a set bijection between $J$ and $\cP$ with the property that for each $j\in J$ the propagator associated to $j$ under the bijection is supported on vertex $j$.

This is equivalent to the condition for a set to be independent or a basis established in Lemma 3.5 and Theorem 3.6 of \cite{wilsonloop}. The equivalence can be seen by applications of the the pigeonhole principle in one direction, and inductively in the other. Below, we provide a different proof of this condition for independence. \hlfix{Note that either definition of independence and bases corresponds to the sets independent columns and bases of $C(W)$.}{I'm having trouble parsing this sentence: help?}

\begin{proof} \sanote{ I would still argue that this should be moved to section 2. That is the only place we talk about matroids and positroid associated to Wilson loop diagrams. While this is a general linear algebra result, it does reference $C(W)$ several times in the proof.}
Because the nonzero entries of $C(W)$ are independent indeterminants, $J$ is an independent set is if and only if there is some choice of $|J|$ nonzero entries of $C(W)$ one in each row associated to an element of $J$ and each in different columns.

Each entry in $C(W)$ identifies a propagator by the row of the entry and a vertex by the column of the entry.  The entry is nonzero if and only if the propagator is supported on that vertex.

Consequently, a choice of $|J|$ nonzero entries of $C(W)$ one in each row associated to an element of $J$ and each in different columns is equivalent to an assignment of the propagators of $J$ to supporting vertices so that no two are assigned to the same vertex.  Such an assignment of the propagators of $J$ to supporting vertices is exactly a map $f$ as described in the statment, hence proving the result.
\end{proof}

@@

We have shown that $(I_1,\dots,I_n)$ is a Grassmann necklace; it remains to check that this Grassmann necklace corresponds to the positroid $M(W)$.  {\color{violet} By equation \eqref{basesofmatroids}, w}e need to show that: \sanote{added a number to this eq. make it a dfn?}
\begin{itemize}
\item For each $i \in [n]$, $I_i$ is a basis for $M(W)$.
\item If $J$ is lexicographically smaller than {\color{violet} any} $I_i$ with respect to $<_i$, then $J$ is not a basis for $M(W)$. 
\end{itemize}
The first point follows immediately: since the algorithm is pairing each $j \in I_i$ with a unique propagator supported on that vertex, $I_i$ is a basis for $M(W)$ by Lemma~\ref{lem basis as perm}.

Now suppose we have a $k$-set $J$ such that $J$ is a basis for $M(W)$ and yet is lexicographically less than $I_i$ with respect to $<_i$.  By Lemma~\ref{lem basis as perm} there is a set bijection between $J$ and the propagators of $W$ such that for each $j \in J$, the propagator associated to $j$ is supported on vertex $j$.  Choose one such bijection.  For each propagator $p$ of $W$, write $J(p)$ for the vertex $j$ associated to $p$ according to this bijection.

@@

\begin{proof}
\begin{enumerate}
    \item The nonzero entries of $C(W)$ are independent indeterminates and so every $i\times i$ minor of $C(W)$ is either homogeneous of degree $i$ or is $0$.  Thus each $\Delta_{I_i}$ and each $r_i$ is homogeneous.  %Furthermore, each row contributes one factor to each term in the expansion of $\Delta_{I_i}$ so the result of setting the variables from a subset of rows to $1$ is still homogeneous.  Thus each $r_i$ is homogeneous.
    \item Using the expression for the determinant as a sum over permutations from Lemma \ref{lem basis as perm} we see that $\Delta_{I_i}$ is a sum over bijections between $I_i$ and $\mathcal{P}$.  The nonzero terms in this sum are precisely those bijections such that each propagator is associated to one of its supporting vertices in $I_i$, since only those locations in $C(W)$ are nonzero.  Since the nonzero entries of $C(W)$ are independent there can be no cancellation between terms in this expansion.

>>>>

One useful corollary of this, which we will want to keep in mind later, is the following.
\begin{cor}\label{lem basis as perm}
Let $W = (\cP,[n])$ be an admissible Wilson loop diagram and let $M(W)$ be its associated matroid. A subset $J \subseteq [n]$ is an independent set of $M(W)$ if and only if there exists an injective set map $f : J \rightarrow \cP$ with the property that for each $j\in J$ we have $j \in V(f(j))$.
\end{cor}

\begin{proof}
  This follows as a corollary of the previous theorem by induction in one direction and the pigeonhole principle in the other direction.

Forgetting about the matroid and simply thinking about the condition of linear independence in the matrix $C(W)$, the corollary can also be proved directly by linear algebra and the definition of $C(W)$ as follows: because the nonzero entries of $C(W)$ are independent indeterminants, $J$ is an independent set is if and only if there is some choice of $|J|$ nonzero entries of $C(W)$ one in each row associated to an element of $J$ and each in different columns. Each entry in $C(W)$ identifies a propagator by the row of the entry and a vertex by the column of the entry.  The entry is nonzero if and only if the propagator is supported on that vertex.  Consequently, a choice of $|J|$ nonzero entries of $C(W)$ one in each row associated to an element of $J$ and each in different columns is equivalent to an assignment of the propagators of $J$ to supporting vertices so that no two are assigned to the same vertex.  Such an assignment of the propagators of $J$ to supporting vertices is exactly a map $f$ as described in the statment, hence proving the result.
\end{proof}

In particular, this corollary says that a subset $J$ of $[n]$ is a basis of $M(W)$ if and only if there is a set bijection between $J$ and $\cP$ with the property that for each $j\in J$ the propagator associated to $j$ under the bijection is supported on vertex $j$.

@@

We are now ready to prove that Algorithm~\ref{alg:put GN on WLD} does in fact give the Grassmann necklace of the positroid associated to $W$.

@@

We have shown that $(I_1,\dots,I_n)$ is a Grassmann necklace; it remains to check that this Grassmann necklace corresponds to the positroid $M(W)$.  {\color{violet} By equation \eqref{basesofmatroids}, w}e need to show that: \sanote{added a number to this eq. make it a dfn?}
\begin{itemize}
\item For each $i \in [n]$, $I_i$ is a basis for $M(W)$.
\item If $J$ is lexicographically smaller than {\color{violet} any} $I_i$ with respect to $<_i$, then $J$ is not a basis for $M(W)$. 
\end{itemize}
The first point follows immediately: since the algorithm is pairing each $j \in I_i$ with a unique propagator supported on that vertex, $I_i$ is a basis for $M(W)$ by Corollary~\ref{lem basis as perm}.

Now suppose we have a $k$-set $J$ such that $J$ is a basis for $M(W)$ and yet is lexicographically less than $I_i$ with respect to $<_i$.  By Corollary~\ref{lem basis as perm} there is a set bijection between $J$ and the propagators of $W$ such that for each $j \in J$, the propagator associated to $j$ is supported on vertex $j$.  Choose one such bijection.  For each propagator $p$ of $W$, write $J(p)$ for the vertex $j$ associated to $p$ according to this bijection.

@@

\begin{proof}
\begin{enumerate}
    \item The nonzero entries of $C(W)$ are independent indeterminates and so every $i\times i$ minor of $C(W)$ is either homogeneous of degree $i$ or is $0$.  Thus each $\Delta_{I_i}$ and each $r_i$ is homogeneous.  %Furthermore, each row contributes one factor to each term in the expansion of $\Delta_{I_i}$ so the result of setting the variables from a subset of rows to $1$ is still homogeneous.  Thus each $r_i$ is homogeneous.
    \item Using the expression for the determinant as a sum over permutations from Corollary~\ref{lem basis as perm} we see that $\Delta_{I_i}$ is a sum over bijections between $I_i$ and $\mathcal{P}$.  The nonzero terms in this sum are precisely those bijections such that each propagator is associated to one of its supporting vertices in $I_i$, since only those locations in $C(W)$ are nonzero.  Since the nonzero entries of $C(W)$ are independent there can be no cancellation between terms in this expansion.
